### 2020-07-13 codebook.Rmd
Check VIF

---
title: "Codebook created on 2020-07-13 at 2020-07-13 21:50:17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = 'asis')
```

## Dataset description
The data contains 94 cases and 24 variables.

## Codebook

```{r}
summaries <- read.csv("C:/tmp/placebo_adhd/codebook.csv", stringsAsFactors = FALSE)
summaries <- summaries[, !colSums(is.na(summaries)) == nrow(summaries)]
options(knitr.kable.NA = '')
knitr::kable(summaries, row.names = FALSE, digits = 2)
```

### Legend

* __Name__: Variable name
* __type__: Data type of the variable
* __missing__: Proportion of missing values for this variable
* __unique__: Number of unique values
* __mean__: Mean value
* __median__: Median value
* __mode__: Most common value (for categorical variables, this shows the frequency of the most common category)
* **mode_value**: For categorical variables, the value of the most common category
* __sd__: Standard deviation (measure of dispersion for numerical variables
* __v__: Agresti's V (measure of dispersion for categorical variables)
* __min__: Minimum value
* __max__: Maximum value
* __range__: Range between minimum and maximum value
* __skew__: Skewness of the variable
* __skew_2se__: Skewness of the variable divided by 2*SE of the skewness. If this is greater than abs(1), skewness is significant
* __kurt__: Kurtosis (peakedness) of the variable
* __kurt_2se__: Kurtosis of the variable divided by 2*SE of the kurtosis. If this is greater than abs(1), kurtosis is significant.

This codebook was generated using the [Workflow for Open Reproducible Code in Science (WORCS)](https://osf.io/zcvbs/)