### 2020-10-18 manuscript/manuscript.Rmd
Add requested tables and figures

---
title: "Untitled"
output: github_document
date: '`r format(Sys.time(), "%d %B, %Y")`'
bibliography: references.bib
knit: worcs::cite_all
---
<script src="//yihui.org/js/math-code.js"></script>
<!-- Just one possible MathJax CDN below. You may use others. -->
<script async
  src="//mathjax.rstudio.com/latest/MathJax.js?config=TeX-MML-AM_CHTML">
</script>

```{r setup, include=FALSE}
library("worcs")
# We recommend that you prepare your raw data for analysis in 'prepare_data.R',
# and end that file with either open_data(yourdata), or closed_data(yourdata).
# Then, uncomment the line below to load the original or synthetic data
# (whichever is available), to allow anyone to reproduce your code:
# load_data()
knitr::opts_chunk$set(echo = TRUE)
run_everything = TRUE
source("../model_accuracy.R")
```

## Multicolinearity analysis

```{r, message = FALSE}
library(worcs)
library(readxl)
library(tidySEM)
library(metaforest)
# Email from Xavier:
# The dataset has 2 sheets. In sheet #1 you will find the data to be analysed and in sheet #2 the description of the study variables.
load_data()
dictionary <- read.csv("../dictionary.csv", stringsAsFactors = FALSE)

dictionary <- dictionary[!is.na(dictionary$Meaning), 1:2]
dictionary <- dictionary[-c(6,30), ]

# Patient-related covariates: mean age, gender (% of men), illness severity at baseline, previous treatment (treatment naïve), ethnicity (% Caucasian).
#dictionary[7:10, ]
mod_patient <- c(`Mean age` = "mean_age_1",
                 `Male (p)` = "men_1_imputed",
                 `Severity at baseline` = "mean_adhd1_baseline_t_imputed",
                 `Treatment naïve` = "naive_an_crit_incl",
                 `Ethnicity` = "white_1_imputed")

# Intervention related covariates: type of drug, concomitant psychotherapy, legal status of the drug (approved vs non approved (= investigational); approved for ADHD vs non-approved for ADHD)
#dictionary[12:14, ]
mod_intervention <-  c(`Stimulant` = "stimulant_investigated",
                       `Therapy` = "psychother_c",
                       `Approved` = "approved_adhd")

# Study design related covariates: study design (parallel vs cross-over), number of sites (or number of patients per study site or sample size), drug dosage (fixed versus variable), RCT duration, number of study arms, comorbidity as an inclusion criterion, proportion of patients assigned to placebo/probability of receiving placebo, duration of wash-out, % dropouts in the placebo arm, type of analysis (ITT: Yes/No).
#dictionary[16:23, ]
mod_design <- c(`Study design` = "parall_dis",
                `Sites` = "num_centers_imputed",
                `Drug dosage` = "fixed_dosification_imputed",
                `RCT duration` = "length_treat2",
                `Placebo (p)` = "prob_pbo",
                `Comorbidity` = "comorb_incl_crit",
                `ITT analysis` = "itt_analyisis"
)

# Other: Publication year, region (USA/others), sponsorship.
# Risk of bias: High risk of bias using the Cochrane risk of bias tool (Yes/No).
#dictionary[16:23, ]
#data$year <- as.numeric(format(data$publication_date, format = "%Y"))
data$year <- as.numeric(substring(data$publication_date, 1, 4))
mod_other <- c(`Year` = "year",
               `USA` = "usa",
               `Sponsorship` = "pharm_ind_main",
               `Risk of bias` = "high_rob")

mods <- c(mod_patient,
          mod_intervention,
          mod_design,
          mod_other
)

# Code factors
desc <- descriptives(data)
is_factor <- desc$unique == 2 & sapply(data, function(x){all(x %in% c(0,1))})
data[is_factor] <- lapply(data[is_factor], factor)

# Covariates with more than 50% of missing data will be removed.

# According to the protocol, now you should:
#   1.	examine multicollinearity, using the generalized VIF (https://github.com/cjvanlissa/metafor/blob/master/R/vif.rma.r).

dat <- escalc(measure = "MN", mi = data$mean_adhd1_change_1, sdi = data$sd_adhd1_change_1_imputed, ni = data$n_adhd1_change_1, data = data[mods])

res <- rma(yi = dat$yi,
           vi = dat$vi,
           mods = model.matrix(~., data[mods])[,-1])
vifs <- vif(res)
write.csv(vifs, "vif.csv")
worcs:::write_gitig(".gitignore", "!vif.csv")

# Please, let us know what are the findings of this analysis. If no collinearity is found, the next steps will be splitting the sample in two and carrygin out the metaregression/metaforest analyses using the training data set.
```

We examined multicolinearity by running a random effects meta-analysis with all predictors included, and computing the generalized VIF. Values ranged from $[`r report(min(vifs), equals = F)`, `r report(max(vifs), equals = F)`]$, thus indicating no evidence of multicolinearity.

## Split training/testing sample

Subsequently, we split the data into a training and testing sample, using the preregistered seed value.

```{r}
set.seed(4286)
train <- sample(unique(data$id), size = .7*length(unique(data$id)), replace = FALSE)
yaml::write_yaml(train, "split.yml")
train <- yaml::read_yaml("split.yml")
df_train <- dat[data$id %in% train, ]
df_test <- dat[!data$id %in% train, ]
```

The training data had `r nrow(df_train)` cases, and the testing data had `r nrow(df_test)` cases.

## Metaforest model

To build a metaforest model, we first conducted an analysis with random effects weights and default settings, with 20000 iterations to determine the number of trees required to reach convergence.

```{r}
# Run model with many trees to check convergence
set.seed(6473)
check_conv <- MetaForest(yi~.,
                        data = df_train,
                        whichweights = "random",
                        num.trees = 20000)
# Plot convergence trajectory
```

```{r figconverge, echo=FALSE, fig.cap="Convergence plot"}
# Plot convergence trajectory
plot(check_conv)
```

It can be seen that this model has converged within approximately 10000 trees (Figure \@ref(fig:figconverge)). Thus, we will use this number of trees for subsequent analyses.

```{r, echo = FALSE}
# Run recursive preselection, store results in object 'preselect'
set.seed(73829)
mf_rep <- MetaForest(yi~.,
                        data = df_train,
                        whichweights = "random",
                        num.trees = 10000)
preselected <- preselect(mf_rep,
                         replications = 100
                         )
# Plot the results
plot(preselected)
# Retain only moderators with positive variable importance in more than
# 50% of replications
retain_mods <- preselect_vars(preselected, cutoff = .95)

```

Next, we tune the model using the R-package \texttt{caret}, which offers a uniform workflow for any machine learning task. The function \texttt{ModelInfo_mf} tells caret how to tune a MetaForest analysis. As tuning parameters, we consider all three types of weights (uniform, fixed-, and random-effects), number of candidate variables at each split from 2-6, and a minimum node size from 2-6. We select the model with smallest prediction error (RMSE) as final model, based on 10-fold clustered cross-validation. Clustered cross-validation means that effect sizes from the same study are always included in the same fold, to account for the dependency in the data. Note that the number of folds cannot exceed the number of clusters in the data. Moreover, if the number of clusters is very small, one might have to resort to specifying the same number of folds as clusters. Model tuning typically takes a long time; for this small dataset, it might take five minutes; for a very large dataset, it might take hours. However, when your dataset is larger, you will often reach model convergence at a lower number of trees, which reduces computation time.

```{r}
# Load the caret library
if(run_everything){
  library(caret)
# Set up 10-fold CV
cv_folds <- trainControl(method = "cv", 
                           index = createFolds(df_train$yi))

# Set up a tuning grid for the three tuning parameters of MetaForest
tuning_grid <- expand.grid(whichweights = c("random", "fixed", "unif"),
                       mtry = 2:6,
                       min.node.size = 2:6)

# X should contain only retained moderators, clustering variable, and vi
X <- df_train[, c("vi", retain_mods)]

# Train the model
mf_cv <- train(y = df_train$yi,
               x = X,
               method = ModelInfo_mf(), 
               trControl = cv_folds,
               tuneGrid = tuning_grid,
               num.trees = 10000)
  saveRDS(mf_cv, "mf_cv.RData")
}

mf_cv <- readRDS("mf_cv.RData")
best <- metaforest:::best(mf_cv)
# Extract R^2_{cv} for the optimal tuning parameters
r2_cv <- mf_cv$results$Rsquared[which.min(mf_cv$results$RMSE)]

```

Based on the root mean squared error, the best combination of tuning parameters were `r best[1]` weights, with `r best[2]` candidate variables per split, and a minimum of `r best[3]` cases per terminal node. We examined convergence of the final model (Figure \@ref(fig:figconvergetuned)). Then, we examine the $R^2_{oob}$.

```{r}
# For convenience, extract final model
final <- mf_cv$finalModel
# Extract R^2_{oob} from the final model
r2_oob <- final$forest$r.squared
# Examine performance on testing set
test_performance <- model_accuracy(final, 
                                   newdata = data.frame(.outcome = df_test$yi, df_test[mods]),
                                   observed = df_test$yi,
                                   ymean = mean(df_train$yi))
```

```{r figconvergetuned, echo=FALSE, fig.cap="Convergence plot for final model"}
# Plot convergence
plot(final)
```

```{r figvarimp, fig.cap="Variable importance plot"}
# Plot variable importance
VarImpPlot(final)
```

```{r figpd, fig.cap="Marginal relationship of moderators with effect size"}
ordered_vars <- names(final$forest$variable.importance)[
  order(final$forest$variable.importance, decreasing = TRUE)]
# Plot partial dependence
PartialDependence(final, vars = ordered_vars,
                  rawdata = TRUE, pi = .95)
```

We can conclude that the model has converged, and has a positive estimate of explained variance in new data, $R^2_{oob} = `r formatC(r2_oob, digits = 2, format = "f")`$, $R^2_{cv} = `r formatC(r2_cv, digits = 2, format = "f")`$. However, the model has negative explained variance in the testing data, $R^2_{test} = `r formatC(test_performance["r2"], digits = 2, format = "f")`$. Thus, the model appears to be overfit. We will plot the variable importance (Figure \@ref(fig:figvarimp)), and partial dependence plots (Figure \@ref(fig:figpd)).

```{r}
p <- PartialDependence(final, vars = ordered_vars, output = "list", rawdata = TRUE)
source("../get_signs.R")
the_signs <- t(get_signs(p))
# Number variables
p <- lapply(1:length(p), function(i){
  p[[i]]+facet_grid(.~Variable, labeller = labeller(
    Variable = setNames(paste0(i, ". ", ordered_vars[i]), ordered_vars[i])
  ))
})

# Reduce font size
p <- lapply(p, function(x){ x + theme(strip.text.x = element_text(size = 6),
                                      axis.text.x = element_text(size = 6),
                                      axis.text.y = element_text(size = 6))})

p_pdp <- metaforest:::merge_plots(p)
ggsave(
  filename = "rf_pdp_numbered.png",
  p_pdp,
  device = "png")
svg("rf_pdp_numbered.svg")
eval(p_pdp)
dev.off()

dirs <- data.frame(the_signs)
VI <- sort(final$forest$variable.importance, decreasing = TRUE)
VI <- data.frame(Variable = paste0(1:length(VI), ". ", names(VI)), Importance = VI, Direction = dirs$V9[match(names(VI), dirs$V1)])
VI$Variable <- ordered(VI$Variable, levels = rev(VI$Variable))
VI$Direction <- ordered(VI$Direction, levels = c("Positive monotonous", "Mostly positive", "Other", "Mostly negative", "Negative monotonous"))

p_vi <- ggplot(VI, aes_string(y = "Variable", x = "Importance")) + 
  geom_segment(aes_string(x = 0, xend = "Importance", 
                          y = "Variable", yend = "Variable"), colour = "grey50", 
               linetype = 2) +
  geom_vline(xintercept = 0, colour = "grey50", 
             linetype = 1) +
  xlab("Permutation Importance") + 
  theme_bw() + theme(panel.grid.major.x = element_blank(), 
                     panel.grid.minor.x = element_blank(),
                     axis.title.y = element_blank(),
                     axis.text.y = element_text(hjust=0)) + 
  geom_point(aes_string(fill = "Direction"), shape = 21, size = 2) + 
  scale_fill_grey(start = 1, end =0) +
  #scale_fill_manual(values = c("Mostly positive" = "white", "Mostly negative" = "gray80", "Negative" = "black", "Other" = "gray50")) + 
  theme(legend.position = c(0.8, 0.2))
ggsave(
  filename = "varimp.png",
  p_vi,
  device = "png")
svg("varimp.svg")
eval(p_vi)
dev.off()

imp_p <- ranger::importance_pvalues(final$forest, formula = .outcome ~., data = final$data, method = "altmann")
VI$pvalue <- imp_p[,2][match(rownames(VI), rownames(imp_p))]
write.csv(VI, "table4.csv", row.names = FALSE)
metaforest:::predict.MetaForest()
df <- as.data.frame(data[data$id %in% tmp, match(names(final$data)[-1], names(data))])
pred <- predict(final, data = df)
write.table(pred$predictions, "clipboard", sep = "\n")
```